An Open Platform for Managing Short Programming Exercises

### 2013-06-20 paper/paper.tex
various small tweaks

\keywords{CS1, novice programmers, data collection, student behavior,
  intelligent tutoring system}

@@

Recent years have seen an explosion of online services for learning
to program:  Udacity \cite{udacity}, CodeAcademy \cite{codeacademy},
ProgZoo \cite{progzoo}, 
Codingbat \cite{codingbat}, Coursera \cite{coursera}, Turingscraft
\cite{turingscraft}, Problets \cite{Kumar:2005:GPA:1163405.1163408}, CS Circles \cite{Pritchard:2013:CCI:2445196.2445370} and
several others.  These systems are very powerful, and show great promise to
improve introductory computer science.  Given all of these
very good systems, one may ask why we thought it a good career move
to build yet another system.  There are three main reasons:

@@

CloudCoder supports two types of exercises:  {\em method} or {\em function-based}
exercises, and {\em whole-program} exercises.  Function based exercises look
a lot like Codingbat exercises:  Students are given a description of
what the method should do, and a couple of example input/output pairs
similar to a unit test.  Writing a new function-based exercise
requires a description, some ``starter code'' (typically the method
header, although exercises can provide as much or as little code as
desired), and a series of inputs and their corresponding outputs.
Test-cases can be marked as ``secret'' so that students don't see their
outputs to disincentivize students from writing a special case to
recognize each input.

@@

There are a number of questions, both philosophical and research-oriented,
that we would like to address in the context of our future work with CloudCoder.

{\bf Can the adoption of programming exercises lead to improved outcomes
in introductory programming courses?}  This is the most obvious
question.  One scenario we are particularly interested in is whether
the use of programming exercises, when {\em added} to an existing course with
no other changes in pedagogy, lead to improved outcomes.  We feel
that this scenario is important because educators have limited time
and resources available to change the way their courses are taught.
If programming exercises are an easy change to implement, then
if adopted on a wide scale they could have a significant impact.

{\bf Can exercises be reused effectively outside of the course in which
they were originally created?}  When assigning a programming exercise to
students, the instructor can create one from scratch, or use a
previously-written exercise.  Since creating good exercises is labor-intensive,
being able to use a previously-written exercise represents a significant
time savings.  However, finding an existing exercise that is {\em appropriate}
is not necessarily a trivial task.  We are interested in finding
out whether exercises written in the context of one course can be reused
effectively in another.  There are many interesting aspects of
this question:

>>>>

\keywords{CS1, novice programmers, data collection, student behavior}

@@

Recent years have seen an explosion of online services for learning
to program:  Udacity \cite{udacity}, Codecademy \cite{codeacademy},
ProgZoo \cite{progzoo}, 
Codingbat \cite{codingbat}, Coursera \cite{coursera}, Turingscraft
\cite{turingscraft}, Problets \cite{Kumar:2005:GPA:1163405.1163408}, CS Circles \cite{Pritchard:2013:CCI:2445196.2445370} and
several others.  These systems are very powerful, and show great promise to
improve introductory computer science.  Given all of these
very good systems, one may ask why we thought it a good career move
to build yet another system.  There are three main reasons:

@@

CloudCoder supports two types of exercises:  {\em method} or {\em function-based}
exercises, and {\em whole-program} exercises.  Function based exercises look
a lot like Codingbat exercises:  Students are given a description of
what the method should do, and a couple of example input/output pairs
similar to a unit test.  Writing a new function-based exercise
requires a description, some ``starter code'' (typically the method
header, although exercises can provide as much or as little code as
desired), and a series of inputs and their corresponding outputs.
Test-cases can be marked as ``secret'' so that students don't see their
outputs to discourage students from writing a special case to
recognize each input.

@@

There are a number of questions, both research-oriented and philosophical,
that we would like to address in the context of our future work with CloudCoder.

{\bf Can the adoption of programming exercises lead to improved outcomes
in introductory programming courses?}  This is the most obvious
question.  One scenario we are particularly interested in is whether
the use of programming exercises, when {\em added} to an existing course with
no other changes in pedagogy, can result in improved outcomes.  We feel
that this scenario is important because educators have limited time
and resources available to change the way their courses are taught.
If programming exercises are an easy change to implement, then
if adopted on a wide scale they could have a significant impact.

{\bf Can exercises be reused effectively outside of the course in which
they were originally created?}  When assigning a programming exercise to
students, the instructor can create one from scratch, or use a
previously-written exercise.  Since creating good exercises is labor-intensive,
being able to use a previously-written exercise represents a significant
time savings.  However, finding an existing exercise that is {\em appropriate}
is not necessarily a trivial task.  We are interested in finding
out whether exercises written in the context of one course can be reused
effectively in another.
This question has several interesting aspects:
